---
title: "ESM 244 Lab 2"
author: "Sara Orofino"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. Attach packages
```{r}
library(tidyverse)
library(janitor)
library(MASS)
library(ggrepel)
library(RColorBrewer)
library(factoextra)
library(ggbiplot)
library(shinythemes)
```

If you want to override (set) a function manually throughout an entire workspace, you can assign the function explicitly:
```{r}
select <- dplyr::select
```

###1. Ordinal logistic regrression for political party affiliation by ethnicity, age, and sex.

```{r}

pol_party <- read_csv("pol_party.csv") %>% 
  clean_names() %>% #uses janitor package to convert all the column names to snakecase 
  select(participant_id:read_carefully)

# A little more wrangling

pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  filter(birth_year > 1900, political_party != "NA", political_party <= 7) %>% 
  mutate(age = 2016 - birth_year)

# Political party: 1 =  very liberal, 7 = very conservative 

# When doing OLR we want our dependent variable to be of class "ORDERED FACTOR"

# To change the class to an ordered factor:

pol_df$political_party <- factor(pol_df$political_party, ordered = TRUE, levels = c("1", "2", "3", "4", "5", "6", "7"))

# Update our categorical predictor variables to factors too:

pol_df$ethnicity <- factor(pol_df$ethnicity)
pol_df$gender <- factor(pol_df$gender)
```


Basic data exploration and visualization:

```{r}
# For easy data counts use tally; can group by as many variables as you want to get the counts for those groups 

count_eth <- pol_df %>% 
  group_by(ethnicity, political_party) %>% 
  tally()

# Note that the outcomes are so biased to white respondents (can tell in the count_eth scale, only groups of white respondents have at least 15 respondents in each category, the recommended for OLR)

counts_g <- pol_df %>% 
  group_by(gender, political_party) %>% 
  tally()

# Exploratory histogram - "stat = count" can help with factors in your histograms 

ggplot(pol_df, aes(x=political_party)) +
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  scale_fill_brewer(palette = "YlGnBu") +
  facet_wrap(~ethnicity, scales = "free")

ggplot(pol_df, aes(x=political_party)) +
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  scale_fill_brewer(palette = "YlGnBu") +
  facet_wrap(~gender, scales = "free")
```

# THESE DATA ARE PRECARIOUS AND N IS TOO SMALL FOR GROUPS!!!!!!!

```{r}
### THIS IS NOT PUBLISHABLE DATA - very biased and sample sizes are too small 

pol_model <- polr(political_party ~ age + gender + ethnicity, data = pol_df)

summary(pol_model)

# Notes about interpretation - coefficients given are log odds with respect to a reference level 
## Age value is 0.0013 is so tiny that there is almost no influence of age on outcome response 
## Increase in log odds for males relative to the reference level (female)
## Ethnicity response reference level is Asian 

# To look at just the odds instead of log odds:

exp(coef(pol_model))

# If the odds are 1 they have not changed. If they are greater than one, odds have increased, less than 1 odds have decreased 

## Increase in odds of males selecting higher ranks (more conservative) than females of the same age and ethnicity 

#############################################

# Once we've decided on a model, we can always make predicitons using it:

predictions <- predict(pol_model, type = "probs")

df <- data.frame(pol_df, predictions) #To combine dataframe predictions with original data used to create the model 
```

###2. World Bank Environment PCA - using WorldBank environment indicators 

(Check the key to see the wrangling process for the data)

```{r}

wb_10 <- read_csv("wb_10.csv") %>% 
  select(-X1) %>% 
  column_to_rownames('code') # takes column with country code and put it into the row name 

#PCA is specifically for continuous variables - in our dataframe these variables are in columns 3-8

wb_pca <- prcomp(wb_10[3:8], scale = TRUE)

summary(wb_pca)

#PC1 describes ~50% of the data and with PC2 about 75%

# Example biplot - variables are shown as red arrows and countries are shown as points 

ggbiplot(wb_pca, groups = wb_10$region) + #Groups will be shown based on region column in wb_10 df
  geom_text_repel(label=rownames(wb_10), size = 2, segment.size = 0.2) + # used to move the text labels for points
  theme_bw()

# Some example interpretations:
## Clusters are things that are really similar in multivariate space, USA is way to the left on its own 

# factoextra package (built to help visualize outcomes of ordination methods)

#scree plot:

fviz_screeplot(wb_pca)

# See contributions of variables to different components:

fviz_contrib(wb_pca, choice = "var", axes = 2, top =6) # change axes to 1 and 2 to show for both PCs

# Biplot showing points and vectors 
fviz_pca_biplot(wb_pca, 
                col.var = "cos2",
                label = "var", 
                repel = TRUE)

# slightly nicer biplot - col.var = cos2 will show how well represented variables are on the principal component (higher cos2 aka lighter ones are better represented), label = var lables variables (vectors), repel moves the labels 
```


###3. Shiny Apps 

Note: Shiny apps are about precision of brackets and parenthesis...

To make a new Shiny App - go to add button and click Shiny Web App - title it and keep as a single file. 

Remember you can move data by selecting the dataset in your working directory and click "Copy to" then the folder that was created for the app

